What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-07-02 android-stacks.tex
addressing minor changes - reviewer 3

%\noindent\emph{Exception Documentation}
\textbf{IV-Document all exceptions thrown by a
method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective}.

@@

We begun our investigation using the dataset provided by the GHTorrent
project~\cite{Gousi13}, an off-line mirror of the data offered through the
Github API. The Github API data come in two flavours; a streaming data flow lists
events, such as reporting an issue, happening on repositories in real time,
while a static view contains the current state of entities. The data is stored
in unprocessed format, in a MongoDB database, while metadata is extracted and
stored in a relational database. The GHTorrent dataset covers a broad range of
development activities on Github, including issues. The project has been
collecting data since February 2012. Up to Dec 2013, when we queried it to
retrieve our sample, it included  16.836 Java projects (step 1) which overall contained
356.057 issues.\footnote{A Java project is specified in GitHub with a
specific label.}

After obtaining this sample of projects and associated issues, 
the ExceptionMiner tool, developed in this work, performed a per-project 
mining task as follows: for each project, the content of each issue and issue comments 
were analized and all exceptions and exception stack traces were extracted (step 2). 
Since this initial set of Java project is too diverse (it includes libraries, web-based or
stand-alone applications), in the second phase of this study we restrict the
analysis domain to Android projects (step 3). This decision was driven by the
fact that most open source Android applications are known for dealing
with several sources of exceptions due to multi-threaded execution and
interaction with input/output resources; moreover they share a common
well-defined platform and a limited set of libraries.

>>>>

%\noindent\emph{Exception Documentation}
\textbf{ IV-Document all exceptions thrown by a
method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
in some languages such as C

@@

%The GHTorrent dataset covers a broad range of
% development activities on Github, including the all the issues per project. Such data 
%is stored in unprocessed format, in a MongoDB database, while metadata is extracted 
%and stored in a relational database.

The mixed-methods approach is composed by the following steps:

Step1. Retrieving the Dataset to be Analized. We begun our investigation using the dataset 
provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
offered through the Github API.  GHTorrent has been collecting data from Git since 
February 2012. Up to Dec 2013, when we queried GHTorrent, it included 16.836 Java projects 
which overall contained 356.057 issues.\footnote {A Java project is specified in GitHub with a
specific label.} From GHTorrent we retrieved the metadata related to each Java project (e.g., 
repository's names and short descriptions), and the content of every issue and issue comment 
and stored in a relational database.

Step 2. Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
 developed in this work, to extract the exception names and exception stack traces of each
 issue and issue comment - analyzing the information of each project at a time. This initial 
set of analyzed Java projects included libraries, web-based, Android and stand-alone applications.

Step 3. Restricting the Dataset to Android Projects. We select a subset of all Android 
projects available on the dataset to perform a deeper analysis (which combined mined
 information extrated from issues with source-code analysis and manual inspection 
of projects resources described on next steps). This decision was driven by the
fact that most open source Anedroid applications are known for dealing
with several sources of exceptions due to multi-threaded execution and
interaction with input/output resources; moreover they share a common
well-defined platform and a limited set of libraries.
To identify Android projects, we performed a case insensitive search for 
the term \textsf{android} in the repository's names and short descriptions.  The
heuristic filtered 2.542 repositories, from which 589 apps had at least one
issue containing a stack trace. Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
The filtered set consisted of 482 apps, from which approximately 50\% are also
hosted in Google Market Place.

Step 4. Combining Information Mined From Issues with Bytecode-Code Analysis.

Step 5
. Combining Information Mined From Issues with Source-Code Analysis.
For each one of the analyzed projects, we
downloaded the source code and using custom scripts, we extracted their package
hierarchy (recursive list of of Java packages) and their custom exception types
as described next.

Step 5. Combining Information Mined From Issues with Manual Inspection.
